### 2014-11-23 paper_60_tooling.tex

\documentclass[SDSUThesis.tex]{subfiles} 
\begin{document}

% new section
\section{Tooling}

\subsection{SDLC Dashboard}
This maybe needs a name: Project MPI????  The goal of this project is to provide a tool that
given the 5 files will be able to provide a score for the software department.  It should run 
as a web application.

This section will provide information to install and run a tool for calculating MPI for an organization.
\subsubsection{Introduction}
\subsubsection{Overview}
\subsubsection{Installation}
\subsubsection{Database}
maybe MongoDB, maybe PostgreSQL/MySQL
\subsubsection{Webserver}
The software can either run standalone on cloud-hosted platform such as Heroku.  Try using Play Framework 2 with Authenticate.

\subsection{SDLC-AE}

\end{document}