### 2014-11-23 quotes.tex

\documentclass[SDSUThesis.tex]{subfiles} 
\begin{document}

\newpage

% moves everything to the bottom of the page :-)
\vspace*{\fill}

\begin{flushleft}
\textit{``What you have here is small data.''}
\end{flushleft}

\begin{flushright}
Clay Campbell
\end{flushright}

\vspace{1in}

\begin{flushleft}
\textit{``Every day, three times per second, we produce the equivalent of the amount of data that the Library of Congress has in its entire print collection, right? But most of it is like cat videos on YouTube or 13-year-olds exchanging text messages about the next Twilight movie.''}
\end{flushleft}

\begin{flushright}
Nate Silver
\end{flushright}

\end{document}