### 2014-11-23 signature.tex
updates to signature page and bibliography

\documentclass[SDSUThesis.tex]{subfiles} 
\begin{document}

\begin{center}
\yourtitle
\vspace{0.5in}
\end{center}

\par This thesis is approved as a creditable and independent investigation by a candidate for the Master's of Science degree and is acceptable for meeting the thesis requirements for this degree.  Acceptance of this thesis does not imply that the conclusions reached by the candidate are necessarily the conclusions of the major department.

\vspace{1.5in}

\singlespacing

\hspace*{\fill}\underline{$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$}\\
\hspace*{194bp}\youradvisor\\
\hspace*{\fill}Dissertation Advisor \hspace*{47bp}Date

\vspace{1in}

\hspace*{\fill}\underline{$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$}\\
\hspace*{194bp}\departmenthead\\
\hspace*{\fill}Head, Math and Statistics \hspace*{40bp}Date

\vspace{1in}

\hspace*{\fill}\underline{$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$}\\
\hspace*{194bp}Dean, Graduate School \hspace*{54bp}Date
\end{document}