### 2014-11-25 paper_60_tooling.tex
capitalize titles and rearrange  sections

\documentclass[SDSUThesis.tex]{subfiles} 
\begin{document}

% new section
\section{TOOLING}

\subsection{SDLC DASHBOARD}
This maybe needs a name: Project MPI????  The goal of this project is to provide a tool that
given the 5 files will be able to provide a score for the software department.  It should run 
as a web application.

This section will provide information to install and run a tool for calculating MPI for an organization.
\subsubsection{INTRODUCTION}
\subsubsection{OVERVIEW}
\subsubsection{INSTALLATION}
\subsubsection{DATABASE}
maybe MongoDB, maybe PostgreSQL/MySQL
\subsubsection{WEBSERVER}
The software can either run standalone on cloud-hosted platform such as Heroku.  Try using Play Framework 2 with Authenticate.

\end{document}