### 2015-02-16 paper_80_conclusion.tex

\documentclass[SDSUThesis.tex]{subfiles} 
\begin{document}

\section{FUTURE WORK}

Due to the number of issues with surveys.  One area of future work would be indentifying
a general set of questions that would best fit MPI.  This set would have to include
the best number of questions, order of questions and wording of questions.  Therefore,
any new organization would not have to determine their own survey, but rather just use the
predetermined set of questions.  It would even be advantageous to build a software system to
handle the survey distribution and collection.  Ideally, the results would automatically be
inserted into the appropriate database tables for MPI scoring.

% new section
\section{CONCLUSION}

There are many metrics that can be used to evaluate a Software Development Organization (SDO). 
Knowing which metrics to use and what they all mean can be a daunting task.  MPI is a
proposed solution to the difficulty of measuring an SDO.

Upon completion, this work shall identify:
\begin{itemize}
    \item Define \textbf{What} characteristics should be measured for an SDO
    \item Define \textbf{How} to measure those characteristics
    \item Map the relationship or lack of relationship with a Balanced Scorecard
    \item Create a Framework to store the necessary data for MPI
    \item Outline a Process to generate the MPI score
    \item Provide an example MPI score with real data
\end{itemize}

An entire software development organization (SDO) needs to be measured and analyzed properly, not just the development portion. This document has provided an overview of what indicators need to be measured for an SDO, and how those indicators can be combined to form a single number indicating the performance score of a software development organization.  Also, a framework to store this data was discussed.

\end{document}