### 2015-02-23 paper_30_measure.tex
working on Chapter 3

\documentclass[SDSUThesis.tex]{subfiles} 
\begin{document}

\section{EVALUATING AN SDO}

    Anything can be measured \cite{Hubbard2010}.  Thus, an SDO can be 
    measured.  Proper measurement is crucial for improvement because
    without a starting point it is impossible to determine progress.
    Also, consistent reporting is essential for 
    tracking historical performance.  Many methods are used
    to measure and evaluate SDOs.  Some of the common ones
    will be explained in this section.

    \subsection{METRICS}
        A \textit{metric} can be defined as the science of indirect measurement. Metrics
        are very common in the measurement of SDOs.  
        The following are some examples of metrics that can be collected 
        for an SDO. 
        \begin{itemize}
            \item SLOC - The number of Source Lines of Code 
            \item NOM - The Number of Methods per class
            \item Complexity - A numerical measure of the code complexity,
                some common examples are McCabe \cite{McCabe1976} and 
                Halstead \cite{Halstead1977}
            \item Design - The amount of coupling and cohesion present 
                in the software code
            \item Source Code Analysis - Tools that determine whether 
                code adheres to specified set of rules. Common 
                examples are PMD\footnote{PMD is a source code analysis product.  
                It is not an acronym.} and FindBugs$^{TM}$ \cite{PMD, Findbugs}.
        \end{itemize}
        All these metrics are beneficial, but none of them tell the story of the entire
        SDO.  Most of the metrics, as seen in the list above, around an SDO focus on 
        the source code and development phase. 
        Since metrics are indirect, it can be very difficult match SDO performance with
        a metric or series of metrics.  Metrics are great for tracking, but decision making
        based upon metrics is difficult.

    \subsection{INDICATORS}
        Indicators can be crucial measurements
        within any business setting, and an SDO is no exception. Determining
        the correct indicators for an organization can
        be difficult.  Software development organizations
        are no exception. 
        The differences between the indicators will be explored and
        possible measures for each indicator in an SDO will be
        presented.  Indicators are sometimes refered to as 
        performance measures \cite{parmenter2010}.  The four
        types of indicators important to an SDO are shown
        in Table \ref{tab:indicators}.

        \begin{longtable}{@{}l l}
                \toprule%
                 \centering%
                 {\bfseries Performance Measure}
                 & {\bfseries Description}\\

                \cmidrule[0.4pt](r{0.125em}){1-1}%
                \cmidrule[0.4pt](lr{0.125em}){2-2}%
                % \midrule
                \endhead

                RI (Result Indicator) & What has been done?   \\
                \myrowcolour%
                KRI (Key Result Indicator) & How you have done? \\
                PI (Performance Indicator) & What to do? \\
                \myrowcolour%
                KPI (Key Performance Indicator) & How to dramatically increase performance? \\

                \bottomrule

                \caption{SOFTWARE DEFECT SEVERITY LEVELS}
                \label{tab:indicators}
            \end{longtable}

        More information on the indicators can be found
        in chapter 1 of 
        as well as \cite{florac1999,kaplan1992}.

        \subsubsection{RESULT INDICATORS FOR SDO}
            This section will cover the potential result indicators for an SDO.

        \subsubsection{KEY RESULT INDICATORS FOR SDO}

        \subsubsection{PERFORMANCE INDICATORS FOR SDO}
            This section will cover the potential performance indicators for an SDO.

        \subsubsection{KEY PERFORMANCE INDICATORS FOR SDO}

        Antolic in \cite{Antolic2008} demostrates some ways to measure KPIs for the  software development process.

    \subsection{BALANCED SCORECARD}

        A common technique for organizations to measure themselves is a balanced scorecard.
        A \textit{balanced scorecard} must contain the following six
        characteristics.
        \begin{enumerate}
          \item Financial 
          \item Customer Focus
          \item Environment/Community
          \item Internal Process
          \item Employee Satisfaction
          \item Learning and Growth
        \end{enumerate}

        Balanced scorecards are great for displaying the important information about
        an organization.  However, a balanced scorecard does not specify what
        exactly needs to be tracked.  It is also not specific to an SDO and 
        it does not produce a single number.

        This info comes from \cite{parmenter2010}.

    \subsection{PROJECT MANAGEMENT MEASUREMENT}
    According to Putnam and Myers in \cite{Putnam2013}, the 5 core measurements for managing software
    projects are:

    \begin{enumerate}
        \item \textbf{Quantity of function} usually measured in terms of size (such as source lines of code or function points), that ultimately execute on the computer
        \item \textbf{Productivity} as expressed in terms of the functionality produced for the time and effort expended
        \item \textbf{Time} the duration of the project in calendar months
        \item \textbf{Effort} the amount of work expended in person-months
        \item \textbf{Reliability} as expressed in terms of defect rate (or its reciprocal, mean time to defect)
    \end{enumerate}
    However, these measurements don't focus on the entire SDO. They say nothing about the availablity of the software infrastructure or the satisfaction of the users.

    \begin{description}
      \item[Time] This measurement can be hours, days, weeks, or months.  the time increment does not matter as long as it is consistent.
      \item[Effort] This measurement is person-months, person-weeks, or person-days.
      \item[Quality] This measurement is the defect rate (defects per time period).
      \item[Amount of Work] This measurement has the same units as Effort above. 
      \item[Process Productivity] look into this a bit further
      \item[Function Points] find a definition and a comparison with story points
      \item[Story Points] Look in some agile book
    \end{description}

    Chapter 7 of \cite{Putnam2013} contains some equations to calculate and relate
    these measures. Those equations will be studied and evaluated.

    \subsection{MEASURING OTHER ASPECTS OF AN SDO}

        It is important to note that software departments do not just
        develop software.  The department has many other duties
        including: deploying software, installing server hardware/software,
        writing documentation, surveying users, research, innovation,
        education and other common business duties.

    \subsection{COMBINING THE MEASURES}

        How can the PIs, RIs, KRIs, and KPIs be combined to form a single
        value called an MPI (Master Performance Indicator)?  The field
        of sports analytics also has numerous techniques for combining
        multiple indicators to produce a single number \cite{Cervone2014}.

making decisions based upon historical data.

Data Driven Software Engineering not Data Driven Software Development

\end{document}