### 2015-02-27 paper_60_tooling.tex

\documentclass[SDSUThesis.tex]{subfiles} 
\begin{document}

% new section
\section{TOOLING}

% section not used in final dissertation

\end{document}