### 2015-03-06 acknowledgements.tex
added quotes back in

\documentclass[SDSUThesis.tex]{subfiles} 
\begin{document}

\newpage

\begin{center}
ACKNOWLEDGEMENTS\\
\end{center}

I would like to acknowledge the generous support I received from my family. Thank you
to Emily for providing encouragement and the final push to get me to eventually finish.
Thank you to Ainsley, Porter, Trey, and Ryker for always providing a smile.

I would also like to acknowledge numerous other people for providing
guidance and support during the process.
I will only use first names, but you know who you are. Thank you to: Clay, Jess, Chris,  
Bob, Todd, Leslie, and Mike.

\end{document}