### 2015-03-20 acknowledgements.tex
final format check, added latest PDF

\documentclass[SDSUThesis.tex]{subfiles} 
\begin{document}

\newpage

\begin{center}
ACKNOWLEDGEMENTS\\
\end{center}

I would like to acknowledge the generous support I received from my family. Thank you
to Emily for providing encouragement and the final push to get me to eventually finish.
Thank you to Ainsley, Porter, Trey, and Ryker for always providing a smile.

Next, I would like to thank my advisor, Dr. Gary Hatfield, for the encouragement and guidance
along the way.  Also, thank you to Clay, Jess, and Chris for helping me formulate 
the initial idea.

Finally, I would like to thank my Mom for providing a 
final grammatical proof-reading of the dissertation.

\end{document}